### 2017-07-13 fma.tex
update reviewer 2

% -----------------------------------------------
% Template for ISMIR Papers
% 2017 version, based on previous ISMIR templates

% Requirements :
% * 6+n page length maximum
% * 4MB maximum file size
% * Copyright note must appear in the bottom left corner of first page
% * Clearer statement about citing own work in anonymized submission
% (see conference website for additional details)
% -----------------------------------------------

% Anonymity: authors, GitHub links, acknowledgments

\documentclass{article}
\usepackage[T1]{fontenc} \usepackage[utf8]{inputenc}
\usepackage[svgnames]{xcolor}
\usepackage[bookmarks=false,colorlinks=true,allcolors=DarkBlue]{hyperref} % MidnightBlue, NavyBlue, DarkBlue, Maroon
\usepackage{ismir,amsmath}
\usepackage{graphicx} \graphicspath{{figs/}}
\usepackage[font=small,skip=8pt]{caption}
\usepackage{booktabs, threeparttable, tabularx}
\usepackage{enumitem} \setlist{itemsep=0pt,topsep=0pt,parsep=0pt,partopsep=0pt}
\usepackage{scrextend}

\newcommand{\todo}[1]{{\color{red} #1}}

\newcommand{\ntracks}{106,574 }
\newcommand{\nartists}{16,341 }
\newcommand{\nalbums}{14,854 }
\newcommand{\ngenres}{161 }
\newcommand{\tduration}{343 }
\newcommand{\aduration}{278 }
\newcommand{\size}{917 }
\newcommand{\weblink}{https://github.com/mdeff/fma}
%\newcommand{\weblink}{http://164.132.40.229/fma}

\newcommand{\tspacea}{\hspace{0.7em}}
\newcommand{\tspaceb}{\hspace{0.9em}}

\title{FMA: A Dataset For Music Analysis}

\multauthor
{Michaël Defferrard$^\dagger$ \hspace{.6cm} Kirell Benzi$^\dagger$ \hspace{.6cm} Pierre Vandergheynst$^\dagger$ \hspace{.6cm} Xavier Bresson$^\ddagger$}
{
	$^\dagger$LTS2, EPFL, Switzerland \hspace{0.4cm}
	$^\ddagger$SCSE, NTU, Singapore \hspace{0.4cm}
	{\small Work conducted when XB was at EPFL.}\\
	{\tt\small \{michael.defferrard,kirell.benzi,pierre.vandergheynst\}@epfl.ch, xbresson@ntu.edu.sg}
}
\def\authorname{Michaël Defferrard, Kirell Benzi, Pierre Vandergheynst, Xavier Bresson}

\sloppy % please retain sloppy command for improved formatting
\begin{document}
\maketitle

% about 150-200 words.
\begin{abstract}
We introduce the Free Music Archive (FMA), an open and easily accessible dataset suitable for evaluating several tasks in MIR, a field concerned with browsing, searching, and organizing large music collections.
The community's growing interest in feature and end-to-end learning is however restrained by the limited availability of large audio datasets.
The FMA aims to overcome this hurdle by providing \size GiB and \tduration days of Creative Commons-licensed audio from \ntracks tracks from \nartists artists and \nalbums albums, arranged in a hierarchical taxonomy of \ngenres genres.
It provides full-length and high-quality audio, pre-computed features, together with track- and user-level metadata, tags, and free-form text such as biographies.
We propose a train/validation/test split and three subsets: a genre-balanced set of 8,000 tracks from 8 major genres, a genre-unbalanced set of 25,000 tracks from 16 genres, and a 98 GiB version with 30s clips.
This paper describes the dataset and how it was created, discuss some suitable MIR tasks, and evaluates some baselines for genre recognition.
Code, data, and usage examples are available at \url{\weblink}.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction} % and Motivation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% hook
While the development of new mathematical models and algorithms to solve challenging real-world problems is obviously of first importance to any field of research, evaluation and comparison to the existing state-of-the-art is necessary for a technique to be widely adopted by research communities. Such tasks require open benchmark datasets to be reproducible. % to achieve their purpose, which are sometimes challenging to collect and share.
In computer vision, the community has developed established benchmark datasets such as \href{http://yann.lecun.com/exdb/mnist/}{MNIST} \cite{mnist}, \href{https://www.cs.toronto.edu/~kriz/cifar.html}{CIFAR} \cite{cifar}, or \href{http://www.image-net.org}{ImageNet} \cite{imagenet}, which have proved essential to advance the field. The most celebrated example, the \href{http://www.image-net.org/challenges/LSVRC/2012/}{ILSVRC2012 challenge} on an unprecedented ImageNet subset of 1.3M images \cite{imagenet_challenge}, demonstrated the power of deep learning (DL), which won the competition with an 11\% accuracy advantage over the second best \cite{convnet_imagenet}, and enabled incredible achievements in both fields \cite{dl}.

\begin{table}[t]
	\small
	\centering
	\begin{threeparttable}
	\begin{tabular}{l@{ }rrcc}
		\toprule
		dataset\tnote{1} & \#clips & \#artists & year & audio \\
		\midrule
		\href{https://staff.aist.go.jp/m.goto/RWC-MDB/}{RWC} \cite{RWC} & 465 & - & 2001 & yes \\
		\href{http://calab1.ucsd.edu/~datasets/cal500/}{CAL500} \cite{cal500} & 500 & 500 & 2007 & yes \\
		\href{http://mtg.upf.edu/ismir2004/contest/tempoContest/node5.html}{Ballroom} \cite{ballroom} & 698 & - & 2004 & yes \\
		\href{https://marsyasweb.appspot.com/download/data_sets/}{GTZAN} \cite{gtzan} & 1,000 & $\sim300$ & 2002 & yes \\
		\href{http://www.cp.jku.at/datasets/musiclef/}{MusiClef} \cite{musiclef} & 1,355 & 218 & 2012 & yes \\
		\href{https://labrosa.ee.columbia.edu/projects/artistid/}{Artist20} \cite{artist20} & 1,413 & 20 & 2007 & yes \\
		\href{http://ismir2004.ismir.net/genre_contest/}{ISMIR2004} & 1,458 & - & 2004 & yes \\
		\href{http://www-ai.cs.uni-dortmund.de/audio.html}{Homburg} \cite{homburg} & 1,886 & 1,463 & 2005 & yes \\  % GarageBand
		103-Artists \cite{103artists} & 2,445 & 103 & 2005 & yes \\
		\href{http://www.seyerlehner.info/index.php?p=1_3_Download}{Unique} \cite{unique} & 3,115 & 3,115 & 2010 & yes \\
		\href{http://www.seyerlehner.info/index.php?p=1_3_Download}{1517-Artists} \cite{1517artists} & 3,180 & 1,517 & 2008 & yes \\
		\href{http://www.ppgia.pucpr.br/~silla/lmd/}{LMD} \cite{lmd} & 3,227 & - & 2007 & no \\
		\href{http://anasynth.ircam.fr/home/media/ExtendedBallroom}{EBallroom} \cite{extballroom} & 4,180 & - & 2016 & no\tnote{2} \\
		\href{https://labrosa.ee.columbia.edu/projects/musicsim/uspop2002.html}{USPOP} \cite{uspop} & 8,752 & 400 & 2003 & no \\
		\href{http://calab1.ucsd.edu/~datasets/cal10k/}{CAL10k} \cite{cal10k} & 10,271 & 4,597 & 2010 & no \\ % Swat10k
		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{MagnaTagATune} \cite{magnatagatune} & 25,863\tnote{3} & 230 & 2009 & yes\tnote{4} \\ % MagnaTagATune
		\href{http://jmir.sourceforge.net/index_Codaich.html}{Codaich} \cite{codaich} & 26,420 & 1,941 & 2006 & no \\ % paper 20,849
		\bf \href{\weblink}{FMA} & \bf \ntracks & \bf \nartists & \bf 2017 & \bf yes \\
		\href{http://www.omras2.org/}{OMRAS2} \cite{omras} & 152,410 & 6,938 & 2009 & no \\
		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} & 1,000,000 & 44,745 & 2011 & no\tnote{2} \\
		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} & 2,084,320 & - & 2017 & no\tnote{2} \\
		\href{https://acousticbrainz.org}{AcousticBrainz} \cite{acousticbrainz} & 2,508,627\tnote{5} & - & 2017 & no \\
		\bottomrule
	\end{tabular}
	\begin{tablenotes}
		% 103-Artists is also known as DB-L, Extended Ballroom
		\item[1] Names are clickable links to datasets' homepage.
		\item[2] Audio not directly available, can be downloaded from \\ \href{http://www.ballroomdancers.com}{ballroomdancers.com}, \href{https://www.7digital.com}{7digital.com}, \href{https://www.youtube.com}{youtube.com}.
		\item[3] The 25,863 clips are cut from 5,405 songs.
		\item[4] Low quality 16 kHz, 32 kbit/s, mono mp3.
		\item[5] As of 2017-07-10, of which a subset has been linked to genre labels for the \href{http://www.multimediaeval.org/mediaeval2017}{MediaEval 2017} \href{https://multimediaeval.github.io/2017-AcousticBrainz-Genre-Task}{genre task}.
	\end{tablenotes}
	\end{threeparttable}
	\caption{Comparison between FMA and alternative datasets which can be used for genre recognition.}
	\label{tab:datasets}
\end{table}
% Lists of datasets:
% * http://www.audiocontentanalysis.org/data-sets/
% * A Survey of Evaluation in Music Genre Recognition

\begin{table*}[t]
	\centering
	\begin{threeparttable}
		\begin{tabular}{l@{\hspace{2.7em}}r
			@{\hspace{2.7em}}>{$}r<{$}
			@{}>{${}={}$}c@{}
			>{$}l<{$}
			@{\hspace{2.7em}}>{$}c<{$}
			@{\hspace{2.7em}}c@{\hspace{2.7em}}c}
		\toprule
		dataset & \#samples & \multicolumn{3}{c}{dimensionality} & \multicolumn{3}{c}{size} \\
		\cmidrule{6-8}
				&           & \multicolumn{3}{c}{}               & \text{volume} & [GiB] & \#days \\
		\midrule
		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{MagnaTagATune} \cite{magnatagatune} &
			25,863 & 29\cdot16\cdot10^3\cdot1 && 4.6\cdot10^5 & 1.2\cdot10^{10} & 3 & 8.7 \\
		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} \tnote{1} &
			2,084,320 & 10\cdot44\cdot10^3\cdot2 && 8.8\cdot10^5 & 1.8\cdot10^{12} & 776 & 241 \\
		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} \tnote{2} &
			1,000,000 & 49\cdot27\cdot10^3\cdot2 && 2.6\cdot10^6 & 2.6\cdot10^{12} & 625 & 567 \\
		\href{\weblink}{FMA} \tnote{2} &
			\ntracks & \aduration\cdot44\cdot10^3\cdot2 && 2.4\cdot10^7 & 2.6\cdot10^{12} & \size & \tduration \\
		\midrule
		\href{http://yann.lecun.com/exdb/mnist/}{MNIST} \cite{mnist} &
			70,000 & 28\cdot28\cdot1 && 7.8\cdot10^2 & 5.5\cdot10^{7\phantom0} & 0.011 & - \\
		\href{https://www.cs.toronto.edu/~kriz/cifar.html}{CIFAR-10} \cite{cifar} &
			60,000 & 32\cdot32\cdot3 && 3.0\cdot10^3 & 1.8\cdot10^{8\phantom0} & 0.16 & - \\
		\href{http://www.image-net.org/challenges/LSVRC/2012/}{ILSVRC 2012-17} \cite{imagenet_challenge} \tnote{3,4} &
			1,341,167 & 482\cdot415\cdot3 && 6.0\cdot10^5 & 8.0\cdot10^{11} & 100 & - \\
		\href{http://www.image-net.org}{ImageNet} \cite{imagenet} \tnote{3} &
			14,197,122 & 482\cdot415\cdot3 && 6.0\cdot10^5 & 8.5\cdot10^{12} & 1000 & - \\
		\bottomrule
	\end{tabular}
	\begin{tablenotes}
		\item[1] \href{https://support.google.com/youtube/answer/6039860}{YouTube recommended} sample rate for music videos. Size in GiB calculated for mp3 encoded at 320 kbit/s.
		\item[2] Average duration and sample rate. Most clips from \href{https://www.7digital.com}{7digital} are 30 or 60s and 22 or 44kHz \cite{msd_features}.
		\item[3] Average resolutions. Most applications resize or crop to $256\cdot256$ for a scale of $2.6\cdot10^{11}$ and $2.8\cdot10^{12}$.
		\item[4] Object detection / classification task. It is the most used dataset for image classification.
	\end{tablenotes}
	\end{threeparttable}
	\caption{Comparison between the largest public audio and image datasets. Dimensionality is (length [s] $\cdot$ sample rate $\cdot$ \#channels) for audio and (xdim $\cdot$ ydim $\cdot$ \#channels) for images. Volume is the number of samples times the dimensionality. Size is for a directory containing all \texttt{.mp3} or \texttt{.jpg} in GiB, an indication of the quantity of information affected by the size and diversity of the data. The last column is the number of days necessary to listen to the whole available audio.}
	\label{tab:size}
\end{table*}
% Magnatagature dead links: http://tagatune.org/Datasets.html http://tagatune.org/Magnatagatune.html
% Announcement: https://musicmachinery.com/2009/04/01/magnatagatune-a-new-research-data-set-for-mir/
% Robin: Alors il télécharge aussi des clips audios de 10 secondes, non-compressé et à 48 kHz pour un total de 4 Terabytes! Donc environ 8x la taille de ton dataset il semble.
% uncompressed: 2084320*10*48e3*2*2 / 1024**4 = 3.6
% 320 kbit/s mp3: 2084320*10*320e3 / 8 / 1024**3 = 776

% Xavier: De meme pour Table 2, c'est interessant bien sur, mais encore une fois, les datasets de MSD et autres n'ont pas disponible (directement et facilement) les raw audio data. Cela rend confusing le main message. Aussi, je ne parlerais pas des MNIST, CIFAR et autres image datasets, ce n'est pas le point de ce papier, targeted music, et cela peut rendre notre message confusing.

\begin{figure*}
	\begin{minipage}{0.74\linewidth}
		\centering
		\includegraphics[width=\linewidth]{growth.pdf}
	\end{minipage} \hfill
	\begin{minipage}{0.24\linewidth}
		\centering
		\includegraphics[width=\linewidth]{album_release_year.pdf}
	\end{minipage}
	\caption{(left) Growth of the archive, created in 11/2008. (right) Number of albums released per year (min 1902, max 2017).}
	\label{fig:growth}
	\label{fig:album_release_year}
\end{figure*}

\begin{table*}
	\tiny
	\centering
\begin{tabular}{@{}l@{\tspacea}l@{\tspacea}l@{\tspacea}l@{\tspacea}r@{\tspacea}r@{\tspacea}l@{\tspacea}r@{\tspacea}l@{\tspacea}l@{\tspacea}l@{}}
	\toprule
	\multicolumn{6}{@{}l}{track} & \multicolumn{3}{@{}l}{album} & \multicolumn{2}{@{}l}{artist} \\
	track\_id & title & genres\_all & genre\_top & dur. & listens & title & listens & tags & name & location \\
	\midrule
	150073 & Welcome to Asia & [2, 79] & International & 81 & 683 & Reprise & 4091 & [world music, dubtronica, fusion] & DubRaJah & Russia \\
	140943 & Sleepless Nights & [322, 5] & Classical & 246 & 1777 & Creative Commons Vol. 7 & 28900 & [classical, alternate, soundtrack, piano, ... & Dexter Britain & United Kingdom \\
	64604 & i dont want to die alone & [32, 38, 456] & Experimental & 138 & 830 & Summer Gut String & 7408 & [improvised, minimalist, noise, ... & Buildings and Mountains & Oneonta, NY \\
	23500 & A Life In A Day & [236, 286, 15] & Electronic & 264 & 1149 & A Life in a Day & 6691 & [idm, candlestick, romanian, candle, ... & Candlestickmaker & Romania \\
	131150 & Yeti-Bo-Betty & [25, 12, 85] & Rock & 124 & 183 & No Life After Crypts & 3594 & [richmond, fredericksburg, trash rock, ... & The Crypts! & Fredericksburg \\
	\bottomrule
	\end{tabular}
	\caption{Some rows and columns of the metadata table, stored in \texttt{tracks.csv}.}
	\label{tab:tracks}
\end{table*}

\begin{table}
	\small
	\centering
	\begin{tabular}{|r@{ }l@{\hspace{.8em}}r@{ }l@{\hspace{.8em}}r@{ }l|}
		\hline
		100\% & track\_id     & 100\% & title          &  93\% & number \\
		  2\% & information   &  14\% & language\_code & 100\% & license \\
		  4\% & composer      &   1\% & publisher      &   1\% & lyricist \\
		 98\% & genres        &  98\% & genres\_all    &  47\% & genre\_top \\
		100\% & duration      & 100\% & bit\_rate      & 100\% & interest \\ %samplerate \\
		100\% & \#listens     &   2\% & \#comments     &  61\% & \#favorites \\
		100\% & date\_created &   6\% & date\_recorded &  22\% & tags\\
		\hline
		100\% & album\_id     & 100\% & title          &     &  \\
		 94\% & type          &  96\% & \#tracks       &     &  \\
		 76\% & information   &  16\% & engineer       &  18\% & producer \\
		 97\% & \#listens     &  12\% & \#comments     &  38\% & \#favorites \\
		 97\% & date\_created &  64\% & date\_released &  18\% & tags \\
		\hline
		100\% & artist\_id    & 100\% & name           &  25\% & members \\
		 38\% & bio           &   5\% & \multicolumn{2}{@{}l}{associated\_labels} &  \\
		 43\% & website       &   2\% & wikipedia\_page & &  \\
		& & 5\% & \multicolumn{2}{@{}l}{related\_projects} & \\
		 37\% & location      &  23\% & longitude      &  23\% & latitude \\
		 11\% & \#comments    &  48\% & \#favorites    &  10\% & tags \\
		 99\% & date\_created &   8\% & active\_year\_begin & & \\
		& & 2\% & \multicolumn{2}{@{}l}{active\_year\_end} & \\
		\hline
	\end{tabular}
	\caption{List of available per-track, per-album and per-artist metadata, i.e.\ the columns of \texttt{tracks.csv}. Percentages indicate coverage over all tracks, albums, and artists.}
	%Not all of them are filled for each track.\footnote{See }
	%\texttt{albums.csv} and \texttt{artists.csv}.}
	% genres.csv ?
	\label{tab:metadata}
\end{table}

An alternative to open benchmarking is the approach taken by evaluation challenges such as the music information retrieval evaluation exchange (MIREX) \cite{mirex}. The imposed constraint that input data cannot be made available to participants however incurs an approximately linear cost with the number of submissions as all algorithms have to be run by the organizers. By releasing this open dataset, we realize part of the vision of McFee et al.\ in ''a distributed, community-centric paradigm for system evaluation, built upon the principles of openness, transparency, and reproducibility`` \cite{mirex_critic}.

% {\bf Available music datasets and their limitations.}
Unlike other modalities such as images or text for which a wealth of content is available, the lack of a large-scale, complete and easily available dataset for MIR has hindered research on data-heavy models such as DL.
\tabref{tab:datasets} lists the most common datasets used for content-based MIR.
GTZAN \cite{gtzan}, a collection of 1000 clips from 10 genres, was the first publicly available benchmark MGR dataset. As a result, despites its flaws, it continues to be the most used dataset for MGR \cite{mgr_eval_2}. In addition to its small size, the main problems of GTZAN is mislabelling, repetitions, and distortions, as well as missing metadata (no artist name and song title) which prevent researchers to control for artists or album effects. % Each song is represented by a 22,050Hz Mono 16-bit wav audio file of 30sec.
Looking at \tabref{tab:datasets}, the well-known MagnaTagATune \cite{magnatagatune} and the Million Song Dataset (MSD) \cite{msd} as well as the newer AudioSet \cite{audioset} and AcousticBrainz \cite{acousticbrainz} appear as contenders for a large-scale reference dataset. % They however each have limitations.
MagnaTagATune, which was collected from the \href{https://magnatune.com/}{Magnatune} label and tagged using the \href{http://tagatune.org/}{TagATune} game, includes metadata, audio features and raw audio. The poor audio quality and limited number of songs does however limit its usage.
MSD and AudioSet, while very large, force researchers to download audio clips from online services. AcousticBrainz's approach to the copyright issue is to ask the community to upload music descriptors of their tracks. Although it is the largest database to date, it will never distribute audio.
% Finally, Echonest reassigned the indexing of the MDS songs with its database, which henceworth makes (almost) impossible the use of the most music recent features developed by Echonest.
% Each song data is composed of high-level and medium-level audio features provided by Echonest service and metadata like artist name, song title, track genre, etc. It has its own dedicated website with related code and additional datasets.
% 2.1 million annotated 10-second sound clips associated with 527 classes. The ontology is quite broad and covers sounds from music to cars, engines or animals. 
On the other hand, the proposed dataset offers the following qualities, which in our view are essential for a reference benchmark.

% Value proposition.
\textbf{Avoiding biases.} sampling bias TODO: reviewer 1 comment 15 \cite{mgr_eval_2}

\textbf{Large scale.} Large datasets are needed to avoid overtraining and to effectively learn models that incorporate the ambiguities and inconsistencies that one finds with musical categories. They are also more diverse and allows to average out annotation noise.
While FMA features less clips than MSD or AudioSet, every other dataset with readily available quality audio are two orders of magnitude smaller (\tabref{tab:datasets}).
For statistical learning algorithms, the scale of FMA is comparable to ImageNet and the most used ILSVRC (\tabref{tab:size}) and shall thus be sufficient for the foreseable future.
Note that while image datasets are larger, the amount of data is similar because of their smaller dimensionality.

\textbf{Permissive licensing.} MIR research has historically suffered from the lack of publicly available benchmark datasets, which stem from the commercial interest in music by record labels, and therefore imposed rigid copyright.
The FMA embraced the sole solution to that problem, that is to aim for tracks which license permits redistribution, such as Creative Commons.
All metadata and code produced by ourselves are licensed under the \href{https://creativecommons.org/licenses/by/4.0)}{
%Creative Commons Attribution 4.0 International License (CC BY 4.0)
CC BY 4.0} and MIT licenses.
% TODO paper CC BY 4.0 if published at ISMIR
% TODO licensing
% Creative Commons: MagnaTagATune
% Proprietary, non-redistributable: MSD, AudioSet?

\textbf{High-quality audio.}
\tabref{tab:datasets} shows that while the smaller datasets are usually distributed with audio, often regardless of copyrights, most of the larger do not.
They either contain features derived from the audio or provide links to download the audio from an online service. The problem with the former is that researchers are stuck with the set of chosen features and are prevented to leverage feature learning algorithms or end-to-end learning systems such as DL, while the later makes no assurance that the files or services will not disappear or change without notice.
Moreover, researchers are wary of proprietary features such as those computed by commercial services like \href{http://the.echonest.com/}{Echonest}. They have e.g.\ extracted features by themselves on the MSD by downloading sample audio, a tedious process \cite{msd_features}.
Finally, distributed or downloadable audio are usually clips of 10 to 30 seconds, sometimes of low quality, e.g.\ 32 kbit/s for MagnaTagATune or an average of 104 kbit/s for MSD \cite{msd_features}.
In comparison, FMA comes with CC-licensed, full-length, high-quality and non-modified audio.
% \todo{See Table ? for a quantitative analysis of quality.}

\textbf{Metadata rich.} \todo{redundant with content} In addition to audio, the dataset comes with a wealth of metadata, shown in \tabref{tab:metadata}. All the metadata available through the API has been archived. In addition to the basics like title, album or artist names, this includes per-track genres and user data such as per-track/album/artist favorites, play counts and comments.
% which are necessary to evaluate content-based recommender systems.
There is also free-form text such as per-track/album/artist tags, album description and artist biography. To compare, the MSD only provides artist-level metadata and tags \cite{msd}, while GTZAN offers none at all.
% social such as artist location or website
% images for tracks, albums and artists

\textbf{Easily accessible.} Working with the dataset only requires to download some \texttt{.zip} archives containing \texttt{.csv} metadata and \texttt{.mp3} audio. No need to crawl the web and circumvent rate limits or access restrictions. Besides, we provide
% Todo an helper \href{https://pypi.python.org/pypi/freemusicarchive}{Python package} and
some usage examples in the \texttt{usage.ipynb} Jupyter notebook to start using the data quickly.

\textbf{Future proof and reproducible.} All files and archives are checksummed and hosted in a long-term digital archive.
%\todo{have a DOI and are hosted on \href{https://zenodo.org}{zenodo.org}, a long-term digital archive powered by the \href{https://home.cern}{CERN}}.
Doing so alleviates the risks of songs to become unavailable, to change without notice or for online services to shut down.
Moreover, we share all the code used to (i) collect the data, (ii) analyze it, (iii) generate the subsets and splits, (iv) compute the features and (v) test the baselines. The developed code can serve as a starting point for researchers to compute their own features or evaluate their methods.
Finally, anybody can recreate or extend the collection, thanks to public songs and APIs.

% Western / commercial music, how to qualify broadness?

% Do not present like this because we might want representativeness. Which would necessarily include commercial music.
%{\bf What would be the requirements for an ideal music dataset?} \\
%(1) Availability of raw audio tracks like 30sec or more,\\
%(2) Availability of metadata such as genres, artist, title, year, lyrics, users' ratings, users' comments, etc,\\
%(3) Large-scale dataset to be able to sample the high-dimensional distribution of music genres\\
%(4) Free and easy access of the dataset,\\
%(5) Legality of the dataset, no copyright issue with e.g.\ open access licence.\\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Dataset} % The dataset, dataset, Introducing the FMA dataset
% Describe what it is, what are the metadata and features.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{duration_distribution.pdf}
	\caption{Track duration (min 0, max 3 hours).}
	\label{fig:duration_distribution}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{listens_distribution.pdf}
	\caption{Album listens (min 0, max 3.6 millions).}
	\label{fig:listens_distribution}
\end{figure}

\subsection{The Free Music Archive}

The dataset is a dump of the \href{https://freemusicarchive.org/}{Free Music Archive}, an interactive library directed by \href{https://wfmu.org/}{WFMU}, the longest-running freeform radio station in the United States.
%The archive is funded by the New York State Music Fund, the MacArthur Foundation, the National Endowment for the Arts, and from the project's users.
The website provides a large catalog of artists and high-quality tracks, hand-picked by established audio curators. Each track is legally free to download as artists decided to release their works under a permissive license.
%Cheyenne Hohman, the director of the FMA, \href{http://freemusicarchive.org/member/cheyenne_h/blog/100000_SONGS}{announced} in July 2016 that the website had reached the landmark of 100,000 tracks.
Inspired by \href{https://creativecommons.org/}{Creative Commons} and the open-source software movement, the FMA provides a legal and technological platform for curators, artists, and listeners to harness the potential of music sharing. While the archive is free and open to anyone, written and audio content is curated, and permission to upload or edit content is granted on an invitation basis.\footnote{\href{http://rhizome.org/editorial/2009/may/1/interview-with-jason-sigal-of-the-free-music-archi}{Interview with Jason Sigal of the Free Music Archive, Rhizome.}}

\subsection{Creation} % Collection, Creation, Generation
% How it was collected.

% Collection
As of April 1st 2017, when the dataset was gathered, the online archive largest track id was 155,320, of which 109,727 were valid (the missing 45,594 ids probably correspond to deleted tracks). \figref{fig:growth} shows the growth of the dataset. In addition to per-track metadata, the used hierarchy of 161 genres and extended album (480 not found) and artist (250 not found) information were collected via the available \href{https://freemusicarchive.org/api}{API}.\footnote{See \texttt{webapi.ipynb} to query the API with our helpers.} Finally, mp3 audio was downloaded over HTTPS.
% Cleaning
Out of all collected track ids, 180 mp3s could not be downloaded, 286 could not be trimmed by ffmpeg, and features could not be extracted from 71. Finally, the license of 2,616 tracks prohibited their redistribution, leaving us with \ntracks tracks.

While it may be argued that the dataset should be cleaned, we wanted it to resemble real world data. As such, we did not remove tracks which have too many genres, are too long, belong to rare genres, etc. Moreover, it is hard to set a threshold and the small number of outliers will not impact much the prediction accuracy. Researchers are obviously free to discard any track for training.

%\subsection{Data Cleaning}
% classes with less than 1% of the number of tracks of the biggest class Pop/Rock are removed

%At this stage the number of genres is 138, but half have only a few samples. Thus we decided to only retain the genres with more than 100 samples. This defines a dataset of 77,643 music data with 68 genres. We further cleaned the dataset by removing tracks which belong to ``non-standard'' music genres like 'Noise', 'Garage', 'Sound Collage', 'Singer', 'Audio Collage', 'Glitch', 'Unclassifiable', 'Lo-Fi', 'Spoken', 'Poetry', 'Talk Radio', 'Avant-Garde', 'Experimental', 'Ambient', and 'Field Recording'. We also suppressed the songs with the title 'Untitled'. Besides, we decided to cut the long tail of the power law distribution by removing songs of genres with less than 300 songs.

\begin{table}
	\small
	\centering
	\begin{tabular}{ccclr}
		\toprule
		id & parent & top\_level & title & \#tracks \\
		\midrule
		38 & None & 38 & Experimental & 38,154 \\
		15 & None & 15 & Electronic & 34,413 \\
		12 & None & 12 & Rock & 32,923 \\
		1235 & None & 1235 & Instrumental & 14,938 \\
		%10 & None & 10 & Pop & 13,845 \\
		%17 & None & 17 & Folk & 12,706 \\
		\midrule
		25 & 12 & 12 & Punk & 9,261 \\
		89 & 25 & 12 & Post-Punk & 1,858 \\
		1  & 38 & 38 & Avant-Garde & 8,693 \\
		\bottomrule
	\end{tabular}
	\caption{An excerpt of the genre hierarchy, stored in \texttt{genres.csv}. Some of the 16 top-level genres appear in the top part, while some second- and third-level genres appear in the bottom part.}
	\label{tab:genres}
\end{table}

\begin{figure}
	\begin{minipage}[t]{0.44\linewidth}
		\centering
		\includegraphics[width=\linewidth]{genre_hierarchy.pdf}
	\end{minipage} \hfill
	\begin{minipage}[t]{0.55\linewidth}
		\centering
		\includegraphics[width=\linewidth]{genres_per_track.pdf}
	\end{minipage}
	\caption{(left) Example of genre hierarchy for the top-level Soul-RnB genre. Left number is the \texttt{genre\_id}, right number is the number of tracks per genre. (right) Number of genres per track.}
	\label{fig:genre_hierarchy}
	\label{fig:genres_per_track}
\end{figure}

\subsection{Content}
% Figures and tables from analysis.ipynb
% TODO audio properties: samplerate (% 44kHz), bitrate (% 128, VBR), channels (% mono, stereo)

The metadata collected from the archive\footnote{\texttt{raw\_tracks.csv}, \texttt{raw\_albums.csv}, \texttt{raw\_artists.csv}} was cleaned, uniformly formatted, merged and stored in \texttt{tracks.csv}\footnote{\label{creation}See \texttt{creation.ipynb} for the code which created the dataset.} which \tabref{tab:tracks} shows an excerpt. That file is a relational table where each row represents a track and columns are listed in \tabref{tab:metadata}. All the metadata has been produced by the artists when uploading their music. While the content is curated, the curators focus on the musical content not the metadata.
Metadata includes song title, album, artist and genres; user data like play counts, favorites, and comments; free-form text like description, biography, and tags. Coverage varies across fields, see \tabref{tab:metadata}.
For ease of use, we kept all the metadata in a single table despite the redundancy incurred by e.g.\ the fact that all tracks from a given artist share all artist related columns. From a practical point of view, the problem is mitigated by compression for storage and those columns can be treated as \textit{categorical variables} to minimize memory usage.

\todo{describe the metadata}
In the FMA, tags can be found at the track, album and artist level: 23,496 tracks (22\%), 2,659 albums (18\%) and 1,605 artists (10\%) have at least one tag. \figref{fig:tag_distribution} shows their distribution.\footnote{One of the artist tag is often the artist name. We thus subtracted one.}

The audio for each track is stored in a file which name is the track id. All tracks are mp3-encoded, most of them with sampling rate of 44,100 Hz, bit rate 320 kbit/s (263 kbit/s on average), and in stereo. Figures~\ref{fig:album_release_year},~\ref{fig:duration_distribution} and \ref{fig:listens_distribution} show the distribution of albums per year, track durations, and play counts per album.
See the \texttt{analysis.ipynb} notebook for a much more detailed analysis of the content.
%todo: graph for audio quality

\subsection{Genres}

\todo{The high coverage of genre information and fine-graine... makes the FMA well suited for MGR.}
Labeling noise is unavoidable
The FMA is especially suited for MGR as it features fine genre information, i.e.\ multiple (sub-)genres associated to individual tracks, has a built-in genre hierarchy (\tabref{tab:genres}), and is annotated by the artists themselves. While the artists are the best placed to judge the positioning of their creations, they might be inconsistent and motivated by factors not necessarily objective, such as achieving a higher play count. Ideally those labels should be one of many ground truths, to be complemented by crowd-sourcing and experts (from different music metadata websites).

While there is no agreement on a taxonomy of genres \cite{mir_review_genre}, we followed the hierarchy used by the archive, which is the one the authors had in mind when annotating their tracks. That hierarchy is composed of \ngenres genres of which 16 are roots, the others being sub-genres. \tabref{tab:genres} shows an excerpt of that information along with the number of tracks per genre and the associated top-level genre, that is the root of the genre tree. \figref{fig:genre_hierarchy} shows an excerpt of the tree.
\todo{3 genres limit}

In the per-track table, the \texttt{genres} column contain the genre ids indicated by the artist. Then, given such hierarchical information, we constructed a \texttt{genres\_all} column which contains all the genres encountered when traversing the tree from the indicated genres to the roots. Finally, if the traversal yields only one root, the track is given a top-level genre stored in the \texttt{genre\_top} column.
\figref{fig:genres_per_track}~and~\ref{fig:genre_distribution} shows the number of genres per track and tracks per genre.
%As often encountered in natural categories, it obeys a power law.

While root genre recognition should be treated as a multi-label problem in general, \figref{fig:genres_per_track} shows that half the tracks have a single root genre.

% MSD:
% not strictly genres (e.g.\ one of them is romantic dinner music)
% one non-music category

\begin{figure}[t]
	\centering
	\includegraphics[width=\linewidth]{genre_distribution.pdf}
	\\ \vspace{1em}
	\includegraphics[width=\linewidth]{genre_top_distribution.pdf}
	\caption{(top) Tracks per genre on the full set (min 1, max 38,154). (bottom) Tracks per all top genre on the medium subset (min 21, max 7,103). Note how experimental music is much less represented in the curated medium subset.}
	\label{fig:genre_distribution}
	\label{fig:genre_top_distribution}
\end{figure}

\subsection{Features} % Feature sets

\tabref{tab:features} lists the features we extracted from the audio with the \href{https://github.com/librosa/librosa}{librosa} Python library, version 0.5.0 \cite{librosa}. These are all the features the library is able to extract.
Each feature set (except zero-crossing rate) is computed on windows of 2048 samples spaced by hops of 512 samples. Seven statistics were then computed over all windows: the mean, standard deviation, skew, kurtosis, median, minimum and maximum.
Those 518 pre-computed features are distributed in \texttt{features.csv} for all tracks.\footnote{See \texttt{features.py} for the code which computed the features.}

In addition,
% a number of descriptive medium-level and high-level audio
features extracted by \href{http://the.echonest.com/}{Echonest} are provided for 13,129 tracks. These features include 8 high-level audio features (acousticness, danceability, energy, instrumentalness, liveness, speechiness, tempo, valence), 5 social features (artist discovery, artist familiarity, artist hotttness, song currency, song hotttness), and 224 MFCC-like temporal features for a number of segments \cite{echonest_temporal}. They are distributed in \texttt{echonest.csv}.
%Echonest audio features are described in Kirell's paper
%Those features can be obtained either by uploading the songs to their website, which are then analyzed and sent back to the user, or simply retrieved from their database if the songs are known to them, i.e.\ they were analyzed before. All of which can be done through an \href{http://developer.echonest.com}{API}. In our case, we only retrieved features from songs which were already in their database.

\subsection{Subsets} \label{sec:subsets}

For the dataset to be useful as a development set or for people with lower computational resources, we propose the following sets, each of which is a subset of the larger set:
\begin{enumerate}
	\item \textbf{Full}: the complete dataset, described above. All \ngenres genres, unbalanced with 1 to 38,154 tracks per genre (\figref{fig:genre_distribution}) and 0 to 31 genres per track (\figref{fig:genres_per_track}).
	\item \textbf{Large}: the full dataset with audio limited to 30 seconds clips extracted from the middle of the tracks (or entire track if shorter than 30 seconds). That trimming reduces the size of the data by a factor 10.
	\item \textbf{Medium}: 25,000 of the most popular 30s clips with good metadata and quality audio. Genre unbalanced with 21 to 7,103 clips per top genre (\figref{fig:genre_top_distribution}), but only one of the 16 top genres per clip.
		\todo{This subset was constructed for single root genre prediction. The clips were selected based on ... Non-musical genres, experimental?}
	\item \textbf{Small}: 8,000 30s clips from 8 top genres, balanced with 1,000 clips per genre, 1 genre per clip. This subset is similar to the very popular GTZAN \cite{gtzan} with the added benefits of the FMA, that is metadata, pre-computed features, and copyright-free audio. \todo{The same process was used to select the top 1000 tracks per genre. The selected genres are those with more than 1000 tracks in the medium subset.}
\end{enumerate}
\tabref{tab:subsets} highlights the main differentiating factors between the proposed subsets.
%See \texttt{analysis.ipynb} for an analysis of the subsets' content similar to that shown in \secref{sec:content}.

\subsection{Splits}

% PTB 19/3/3, MNIST 60/10, ImageNet 1200/50/100 --> ratio ~6-12
We propose an 80/10/10\% split into training, validation and test sets to make research on the FMA reproducible. Training and validation shall be merged if cross-validation is used instead.
Below are the followed constraints:
% Data was stratified according to the following criteria: ground truth, artist and times filters
\begin{enumerate}
	\item Stratified sampling to preserve the percentage of tracks per genre (important for minority genres).
	\item An artist filter for artists to be part of one set only, thus avoiding any artist and album effect. It has been shown that the use of songs from the same artist in both training and test sets leads to over-optimistic accuracy and may favor some approaches \cite{artist_filters_1, artist_filters_2}.
%All songs from a given artist either appear in train, validation or test, never in two of them.
	%\item Stratified sampling with a time filter, i.e.\ for each genre using the earlier songs in the training set, and the later releases in the validation and test sets.
	%\item \todo{Artists with lots of tracks are balanced across sets, i.e.\ the three sets were alternatively filled in descending order of the count of artist songs. This does not break the above constraint.}
	\item The 2231 tracks without genre label are assigned to the training set as they might be useful as additional training samples for semi-supervised algorithms.\footnote{These are only found in the full and large sets.}
\end{enumerate}
The above constraints are satisfied for all subsets, and a track is assigned to the same split across all of them.\footref{creation} Each root genre is guaranteed to be represented in all splits, but the ratio is only exact for the small subset (800/100/100). The seven smallest sub-genres (less than 20 tracks) are however not guaranteed to appear in all splits of the full and large sets.

\begin{table}
	\small
	\centering
	\begin{tabular}{lrrccc}
		\toprule
		dataset & clips & genres & length & \multicolumn{2}{c}{size} \\
		\cmidrule{5-6}
		        &       &        &  [s]   & [GiB] & \#days \\
		\midrule
		small  &    8,000 &   8 &  30 &  7.4 & 2.8  \\
		medium &   25,000 &  16 &  30 & 23   & 8.7  \\
		large  & \ntracks & 161 &  30 & 98   & 37 \\
		full   & \ntracks & 161 & \aduration & \size & \tduration  \\
		\bottomrule
	\end{tabular}
	\caption{Proposed subsets of the FMA.}
	\label{tab:subsets}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Usage} % proposed usage, tasks, usage, problems
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% List some MIR tasks and propose some who can be done with that dataset

With its rich set of metadata, user data, audio and features, the FMA is amenable to many tasks in MIR. We share below some possible uses which serve to illustrate the breadth of data available in the dataset.

% Metadata analysis
% Recommendation / Recommender systems, Music Similarity
% Music classification
	% Genre recognition / classification
	% Artist identification / recognition
	% Singer / composer recognition
	% Year prediction / recognition
% Automatic (music) tagging / autotagging --> tag prediction

% Segmentation
% Visualization
% Musicology

%\subsection{Album covers}
% image covers: generate covers per genre / artist
% study the influence of covers on ratings

\subsection{Music Classification and Annotation}
%\subsection{Automatic Music Tagging}

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{tag_distribution.pdf}
	\caption{Per-track, album and artist tags (min 0, max 150).}
	\label{fig:tag_distribution}
\end{figure}

Music classification is a key problem in MIR with many potential applications. For one, a classification system enables end users to search for the types of music they are interested in. On the other hand, different music types are managed more effectively and efficiently once they are categorized into different groups \cite{mir_review_classif}.
The classification tasks which can readily be evaluated on FMA include genre recognition, artist identification, year prediction, and automatic tagging.

Automatic tagging \cite{autotagging} is a classification problem which covers different semantic categories, where tags are labels which can be any musical term that describes the genre, mood, instrumentation, and style of the song.
It helps to convert the music retrieval problem to text retrieval by substituting songs with tags.
%The task of music annotation has recently gained much popularity in the MIR community.

In addition to supervised methods which classify music given an arbitrary taxonomy, another approach is to cluster data in an unsupervised way so that a categorization will emerge from the data itself based on objective similarity measures. Then, does genre or another taxonomy naturally come up?

\subsection{Genre Recognition} % MGR, Baselines

Music genres are categories that have arisen through a complex interplay of cultures, artists, and market forces to characterize similarities between compositions and organize music collections. Yet, the boundaries between genres still remain fuzzy, making the problem of music genre recognition (MGR) a nontrivial task \cite{mir_review_genre}.
% genre not defined by audio content only
While its utility has been debated, mostly because of its ambiguity and cultural definition, it is widely used and understood by end-users who find it useful to discuss musical categories \cite{mgr_why}.
As such, it is one of the most researched areas of MIR.
We propose the following problems of increasing difficulty:
\begin{enumerate}
	\item Single top genre prediction on the small subset.
	\item Single top genre prediction on the medium subset.
	\item Multiple top genres prediction on the large / full set.
	\item Multiple genres prediction on the large / full set.
\end{enumerate}

\tabref{tab:mgr} reports accuracies for problem 2 with nine mainstream feature sets and some combinations as well as four standard classifiers using scikit-learn, version 0.18.1 \cite{scikit-learn}. Specifically, we employed linear regression (LR) with an $L^2$ penalty, k-nearest neighbors (kNN) with $k=200$, support vector machines (SVM) with a radial basis function (RBF) kernel and a multilayer perceptron (MLP) with 100 hidden neurons. All classifiers were tested with otherwise default settings.\footnote{See \texttt{baselines.ipynb} for all details.} Reported performance should not be taken as the state-of-the-art but rather as a lower-bound and an indication of the task's difficulty. Moreover, the developed code can serve as a reference and is easily modified to accommodate other features and classifiers.

A major motivation to construct this dataset was to enable the use of the powerful DL set of techniques to music analysis, an hypothesized cause of stagnation on MIREX tasks \cite{mirex_stagnation}.
With availability of audio, DL architectures such as convolutional neural networks and recurrent neural networks can be applied to the waveform to avoid any feature engineering. While those approaches have fallen behind learning from higher-level representations such as spectrograms \cite{dieleman_endtoend}, a greater exploration of the design space will hopefully provide alternatives to solving MIR challenges \cite{mir_dl_feature_learning}.

%Some very recent advances are indeed promising Chuo

\begin{table}
	\small
	\centering
	\begin{tabular}{l@{\tspaceb}r@{\tspaceb}c@{\tspaceb}c@{\tspaceb}c@{\tspaceb}c}
		\toprule
		feature set & dim. & LR & kNN & SVM & MLP \\
		\midrule
		1 Chroma \cite{chroma} & 84 & 44 & 44 & 48 & 49 \\
		2 Tonnetz \cite{tonnetz} & 42 & 40 & 37 & 42 & 41 \\
		3 MFCC \cite{mfcc} & 140 & 58 & 55 & 61 & 53 \\
		4 Spec. centroid & 7 & 42 & 45 & 46 & 48 \\
		5 Spec. bandwidth & 7 & 41 & 45 & 44 & 45\\
		6 Spec. contrast \cite{contrast} & 49 & 51 & 50 & 54 & 53 \\
		7 Spec. rolloff & 7 & 42 & 46 & 48 & 48 \\
		8 RMS energy & 7 & 37 & 39 & 39 & 39 \\ % Root-mean-square energy
		9 Zero-crossing rate & 7 & 42 & 45 & 45 & 46 \\
		\midrule
		% Number and use them for aggregates
		3 + 6 & 189 & 60 & 55 & 63 & 54 \\
		3 + 6 + 4 & 273 & 60 & 55 & 63 & 53 \\
		1 to 9 & 518 & 61 & 52 & 63 & 58 \\
		%\midrule
		%Ech. audio & 8 & - & - & - & - \\
		%Ech. social & 5 & - & - & - & - \\
		%Ech. temporal \cite{echonest_temporal} & 224 & - & - & - & - \\
		\bottomrule
	\end{tabular}
	\caption{Test set accuracies of various features and classifiers for top genre recognition on the medium subset.}
	\label{tab:mgr}
	\label{tab:features}
\end{table}

\subsection{Data Analysis}
% Musicology?
% metadata: better suited datasets. Focus on audio.

While our intention was to release a large volume of audio for machine learning algorithms, analyzing audio is certainly of interest to musicologists and researchers who want to study relations with higher-level representations. Moreover, the metadata is surely a valuable addition to existing datasets\footnote{Like MusicBrainz, AllMusic, Discogs and Last.fm.} for metadata analysis.
%For instance, one could address questions like: How broad artists are, i.e.\ to which genres do they contribute to? Do we observe some drift in time for the vocabulary used in song titles, album or artist names? Which semantic categories can we find in tags and to what extent? How are tags correlated (e.g.\ rock \& guitar, disco \& violin)?
%Other lines of work are to test algorithms for the imputation of missing data.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\subsection{Limitations}
% ground truth for genre: AllMusic Guide (album, artist), Gracenote CDDB (song)
% genre taxonomy: AllMusic, Amazon, mp3.com
% musicbrainz, playme, 7digital
% Allmusic: album reviews, artist biographies, discographies as well as classification of albums according to genres, styles, moods and themes. Information is provided and curated by experts.
% playlists: Art of the Mix playlist categories
% AllMusic and Discogs are based on editorial metadata databases maintained by music experts and enthusiasts 
% Lastfm and Tagtraum are based on collaborative music tagging platforms with large amounts of genre labels provided by their users for music recordings.

% Cover song detection / recognition --> why not, we have track titles
% Lyrics
% Automatic playlist generation
% style?
% Music classification
	% mood classification (happy, sad, and angry) --> may be in tags
	% instrument recognition / identification --> may be in tags
	% music annotation --> temporal labels, e.g.\ instrument / note

% metadata such as genre, mood, groove, composer, performer, lyrics, meter, chord progressions, instruments present

Music classification is the most straightforward use case for FMA, with MGR being particularly suitable. The inclusion of a genre hierarchy makes it specially interesting, as it offers possibilities rarely found in alternative collections.

While the FMA can be used to evaluate many tasks, metadata is missing for e.g.\ mood classification or instrument recognition. However, a more thorough investigation of the available tags may reveal their feasibilities. Similarly, cover song detection may be doable if multiple versions of many songs are featured.
Moreover, we can expect the dataset to be cross-referenced with other resources to unlock additional tasks, as has happened for example with the MSD and \href{http://www.allmusic.com}{AllMusic}, \href{https://www.last.fm}{last.fm} and \href{https://beatunes.com}{beaTunes} for genre recognition \cite{msd_features, msd_genres}, \href{https://musixmatch.com}{musixmatch} for lyrics, \href{https://secondhandsongs.com}{SecondHandSongs} for cover songs, or \href{https://www.thisismyjam.com}{This Is My Jam} for user play counts.

Diversity is another issue: as suggested by \figref{fig:genre_distribution}, this collection is biased toward experimental, electronic, and rock music. Moreover, it probably does not contain hits from renown artists or commercial music.
\todo{more discussion after reading Sturm}
study to what extent is the FMA representative of mainstream music and its similarity to other datasets.
\todo{McFee}
% features some renown artists (wikipedia list)

%The full potential of the FMA has yet to be unlashed.

\subsection{Music Recommendation} % Recommendation, Recommender Systems
% Cite Kirell's paper: Song recommendation with non-negative matrix factorization and graph total variation
%Cheyenne: We do have lists for each user of which songs, albums and artists a user marked as 'favorites' though.
%friends form a social networks
% fans and favorites
% user mixes

Due to the potential commercial value of a working system, music recommendation and music similarity are among the best studied areas of MIR.
%So far, content-based systems have fallen short at predicting user ratings when compared to collaborative filtering methods. \todo{ref}
While the present dump only captures listening and downloading counts in aggregates (\figref{fig:listens_distribution})\footnote{That information can be useful to e.g.\ predict hits.}, the lists of which songs, albums and artists a user marked as favorites or commented are public. While not public, listening and downloading activities are logged and might be shared after anonymization.\footnote{Private discussion with the administrators.} Moreover, users form a public social network via \textit{friend requests}.
%The wealth of available data opens the possibility of a large-scale evaluation of content-based recommender systems based on rich metadata, audio, or derived features.
%with a clean ground truth.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion and Perspectives}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MSD: the future of the dataset, visibility for MIR (to DL people!)

% importance of open data

Benchmarking is an important aspect in experimental sciences -- results reported by individual research groups need to be comparable. Important aspects of these are datasets that can be easily shared among researchers, together with a set of defined tasks and splits. The FMA enables researchers to test algorithms on a large-scale collection, closer to real-world environments. Even though it is still two orders of magnitude behind commercial services who have access to tens of millions of tracks,\footnote{\href{http://the.echonest.com}{37M Echonest}, \href{https://en.wikipedia.org/wiki/Spotify}{30M Spotify}, \href{http://www.skilledtests.com/wiki/Last.fm_statistics}{45M last.fm}, \href{http://bupz.com/best-websites-to-buy-musics}{45M 7digital}, \href{https://www.apple.com/pr/library/2013/02/06iTunes-Store-Sets-New-Record-with-25-Billion-Songs-Sold.html}{26M iTunes}} it is of the same scale as the largest image dataset (\tabref{tab:size}) which opened the door to dramatic performance improvements for many tasks in computer vision.
%is also far behind the amount of tagged photos \href{https://www.facebook.com}{Facebook} has access to for example.
By providing audio, we do not limit the benchmarking to pre-computed features and allow scientists to develop and test new feature sets, learn features, or learn mappings directly from the audio.
% talk about metadata ?
% Kirell: less DL oriented, some people don't like it. Insist on available audio.

% Once that problem is well advanced, researchers will be able to tackle more difficult tasks such as tagging or annotation.
% Besides, standard data analysis tasks such as classification, clustering, regression, text analysis, visualization can also be applied to this new music dataset.

% Future work
\todo{If the community finds interest in the dataset and validate its use, future work should focus on validating the ground truth and collecting additional information.}
In addition to the proposed usage and many others people will find, future work on the dataset itself may consider to scrap the website for information not available through the API, e.g.\ the number of downloads or the user comments. Moreover, cross-referencing with other resources or crowd-sourcing (with e.g.\ \href{https://www.mturk.com}{Mechanical Turk} or \href{https://www.crowdflower.com/}{CrowdFlower}) is desirable to obtain additional labels or measure agreement on the current ones by independent annotators.

In a \href{https://freemusicarchive.org/member/cheyenne_h/blog/FMA_Dataset_for_Researchers}{post about the dataset}, Cheyenne Hohman, the Director at the Archive, wrote that ''by embracing the
~\ldots~%some rights reserved
philosophy of Creative Commons, artists are not only making their music available for the public to listen to, but also for educational and research applications``.
Let's hope for a future where sharing is first and researchers feed open platforms with algorithms while they feed us with data.
% Foster usage of CC and tighter interplay / integration between artists, platforms, users and researchers.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We want to thank the team supporting the \href{https://freemusicarchive.org}{Free Music Archive} as well as all the contributing artists and curators for the fantastic content they made available.
We want to thank the anonymous ISMIR reviewers for their thorough reviews and many constructive comments which have improved the quality of this work.
We are grateful to SWITCH and EPFL for hosting the dataset within the context of the \href{https://projects.switch.ch/scale-up}{SCALE-UP} project, funded in part by the swissuniversities \href{http://www.swissuniversities.ch/isci}{SUC P-2 program}.
% We are grateful to \href{https://zenodo.org}{zenodo.org}, whose free hosting and archiving of digital research artifacts is promoting a more open scientific process.
Xavier Bresson is supported by NRF Fellowship NRFF2017-10.

\bibliography{refs}
\end{document}